Device fingerprinting for user authentication in web applications \titlenote{(Does NOT produce the permission block, copyright information nor page numbering). For use with ACM\_PROC\_ARTICLE-SP.CLS. Supported by ACM.}

### 2018-11-16 paper.tex
Added fingerprint as sole auth method discussion (#6)

\section{Extensions}
\subsection{Improvements to be made}
One of the issues regarding the current fingerprinting technology that we did not cover in depth is the comparison between accuracy and convenience of the technology. As mentioned previously in Section 3.3, in order to use fingerprinting technology as an authentication method, there needs to be stability across different scenarios. Minor changes made to a user's device such as a screen resolution change have to be handled either by accepting or rejecting the changes in the fingerprint. However, the only way that we can accept this change in our current implementation is by ignoring this specific set of data. This is because device fingerprinting results in a hash and any changes can cause the hash to be totally different.

\paragraph{Accuracy vs Convenience}
In order to increase the accuracy of fingerprinting technology, we have to include as many different data that can be acquired from the device as possible so as to get more unique values. This results in less convenience for the users as they have to fall back to a two factor authentication in order for the system to recompute their fingerprint. Further research will have to be conducted to find the set of data that is most optimized for both accuracy as well as convenience. Another method to improve the convenience without affecting too much of the accuracy is to use fuzzy hashing to compute the fingerprint hash as fuzzy hashing allows for small changes in the input without affecting the resulting hash. Therefore, it will then be able to handle changes that are small while still detecting large changes such as logging in from a completely different system.

>>>>

\section{Extensions}

\subsection{Improvements to be made}

In our current implementation, we allow for variations in the components of a fingerprint by assigning a weight to each component that represents the unlikelihood of the value changing. These weights are used during the authentication stage to calculate how different the current fingerprint is compared to the original fingerprint stored in the database. However, in order to support such variations, there are few other issues that we have to consider.

\paragraph{Accuracy or Convenience}
The tradeoff between accuracy and convenience is one such issue that we did not cover in detail. Even though our implementation includes various security levels, we did not investigate the optimal threshold level that should be set for the different levels of security or the optimal algorithm to calculate the amount of variation for each component.

Increasing the threshold for the amount of variation will increase the accuracy of  authentication while at the same time provide inconvenience to the users since they are unable to make any changes to their device configuration. On the other hand, if the threshold is lowered, though the authenticity of the users might be affected, users are instead able to log in successfully even though they have installed a new font or changed their screen resolution after they registered their fingerprints. To aid with the optimal threshold level, custom algorithms for check each component can also be implemented. For instance, a change in the language component from en-US (English - United States) to en-SG (English - Singapore) can have a lower amount of variation compared to a change from en-US to ru-RU (Russian - Russia). Further studies will have to be conducted to find the optimal threshold level, the weights for each component as well as the algorithm to check each component.

Another concern regarding accuracy and convenience is that there is no one security level that is able to fit different use cases. For a certain use case, using a medium security level may be sufficient while for another specific use case, we may have to use a low security level. One method to go about solving this issue is to allow for more detailed customization by the users themselves. For example, if a certain user travels frequently due to work, the user will have the value of the timezone component change regularly. In this case, it is better to reduce the effect the timezone component has on the amount of variations. Therefore, the timezone component can adjusted a lower weight just for this specific user.

\paragraph{Change of Hardware Configurations}
Though most devices connected to the internet are mobile devices such as smartphones, tablets and laptops, where the hardware are unlikely to change, there are still a significant number of desktop computers being used. The hardware of desktop computers can change over time since users will likely upgrade their computer's components as they get outdated. This can reduce the success rate of fingerprint authentication and will require the users to re-register their fingerprint when the new device fingerprint differs too much from the original. Instead of updating the fingerprint every time a variation is detected, a proposed solution is to normalize the device fingerprint over time. This allows for the device to go through different configurations over a period of time while still maintaining both the accuracy and convenience. It is also able to handle the use case mentioned previously where a the timezone component can also change over time when a user is travelling over multiple countries.

\paragraph{Multi-device Support}
Adding multi-device support to our current implementation is as easy as registering multiple device fingerprints for each user account. However, though this is possible, supporting multiple devices for a single user comes with its own limitations such as scalabilty. Having multiple fingerprints for one user means that when logging into the account, the server would need to check the current device fingerprint with the list of fingerprints registered under that specific user. As the size of the fingerprint is larger than regular passwords and each components have to be checked with its own algorithm, a user with a significant number of devices will take a much longer time to authenticate. Further more, because our implementation allows for  variations in the device fingerprints, we will be unable to tell which fingerprints stored on the server is the current fingerprint's variation. This prevents us from using the method where we normalize a fingerprint over time.

\subsection{Real World Implementation}
Before we start using fingerprint authentication in a real world application, it is very important to ensure that a 2FA (Two Factor Authentication) is in place. This is because in the event where a user's device fingerprint changes significantly more than the allowed threshold, we are still able to use 2FA to authenticate the user, and at the same time reregister the user's device fingerprint into the database.
